Cluster efficient pangenome graph construction with nf-core/pangenome

### 2024-06-11 manuscript/main.tex
more feedback
correct name Michael L. Heuer

	\author[Heumos \textit{et~al}.]{
        Simon~Heumos\,$^{\orcidlink{0000-0003-3326-817X}1,2,3,4,*}$,
        Michael~F.~Heuer\,$^{\orcidlink{0000-0002-9052-6000}}$,
        Friederike~Hanssen\,$^{\orcidlink{0009-0001-9875-5262}1,2,3,4}$,
        Lukas~Heumos\,$^{\orcidlink{0000-0002-8937-3457}5,6,7}$,
        Andrea~Guarracino\,$^{\orcidlink{0000-0001-9744-131X}8,9,}$,
        Peter~Heringer\,$^{\orcidlink{0009-0005-5985-2317}1,2,3,4}$,
        Philipp~Ehmele\,$^{\orcidlink{0000-0001-5945-7839}5}$,
        Pjotr Prins\,$^{\orcidlink{0000-0002-8021-9162}8}$,
        Erik~Garrison\,$^{\orcidlink{0000-0003-3821-631X}8}$,
        Sven Nahnsen\,$^{\orcidlink{0000-0002-4375-0691}1,2,3,4,*}$
    }

    \address{
	$^1$Quantitative Biology Center (QBiC) Tübingen, University of Tübingen, Tübingen, Germany \\ 
	$^2$Biomedical Data Science, Dept. of Computer Science, University of Tübingen, Tübingen, Germany \\ 
	$^3$M3 Research Center, University Hospital Tübingen, Tübingen, Germany \\ 
	$^4$Institute for Bioinformatics and Medical Informatics (IBMI), Eberhard-Karls University of Tübingen, Tübingen, Germany \\ 
	$^5$Institute of Computational Biology, Department of Computational Health, Helmholtz Munich, Germany \\ 
	$^6$Comprehensive Pneumology Center with the CPC-M bioArchive, Helmholtz Zentrum Munich, Member of the German Center for Lung Research (DZL), Munich, Germany \\ 
	$^7$TUM School of Life Sciences Weihenstephan, Technical University of Munich, Freising, Germany \\
	$^{8}$Department of Genetics, Genomics and Informatics, University of Tennessee Health Science Center, 71 S Manassas St, Memphis, 38163, Tennessee, USA \\ 
	$^{9}$Human Technopole, Viale Rita Levi-Montalcini 1, 20157, Milan, Italy \\
    }

@@

	Current pangenome graph construction methods exclude complex sequences or are reference-biased \citep{Chin2023, Minkin2016}.
	%, are tree-guided, or reference-based \citep{Li:2020, Hickey2023} leading to reference, order, or orientation bias. 
	%Although whole genome scaling approaches for unbiased pangenome graph construction \citep{Chin2023, Minkin2016} exist, their reliance on k-mer-based data structures often leads to unwanted complexity for downstream analysis. 
	One recent approach that overcomes such limitations is the PanGenome Graph Builder (PGGB) pipeline \citep{Garrison2023}. 
	PGGB iteratively refines an all-to-all whole-genome alignment graph that lets us explore sequence conservation and variation, infer phylogeny, and identify recombination events. 
	PGGB was already extensively evaluated \citep{Garrison2023, Andreace2023} and applied to build the first draft human pangenome reference \citep{Liao2023}.
	However, PGGB is implemented in bash: This (a) makes it difficult to deploy on HPC systems, (b) does not allow for a fine granular tuning of computing resources for different steps of the pipeline \citep{Sztuka2024}, and (c) limits its cluster scalability because PGGB can only use the resources of one node.
	%SN: Given the high resource requirements of large-scale pangenomes, these limitations are significantly limiting broad applications.

@@

	%\vspace{-1.1cm}

>>>>

	\author[Heumos \textit{et~al}.]{
        Simon~Heumos\,$^{\orcidlink{0000-0003-3326-817X}1,2,3,4,*}$,
        Michael~L.~Heuer\,$^{\orcidlink{0000-0002-9052-6000}5}$,
        Friederike~Hanssen\,$^{\orcidlink{0009-0001-9875-5262}1,2,3,4}$,
        Lukas~Heumos\,$^{\orcidlink{0000-0002-8937-3457}6,7,8}$,
        Andrea~Guarracino\,$^{\orcidlink{0000-0001-9744-131X}9,10,}$,
        Peter~Heringer\,$^{\orcidlink{0009-0005-5985-2317}1,2,3,4}$,
        Philipp~Ehmele\,$^{\orcidlink{0000-0001-5945-7839}6}$,
        Pjotr Prins\,$^{\orcidlink{0000-0002-8021-9162}9}$,
        Erik~Garrison\,$^{\orcidlink{0000-0003-3821-631X}9}$,
        Sven Nahnsen\,$^{\orcidlink{0000-0002-4375-0691}1,2,3,4,*}$
    }

    \address{
	$^1$Quantitative Biology Center (QBiC) Tübingen, University of Tübingen, Tübingen, Germany \\ 
	$^2$Biomedical Data Science, Dept. of Computer Science, University of Tübingen, Tübingen, Germany \\ 
	$^3$M3 Research Center, University Hospital Tübingen, Tübingen, Germany \\ 
	$^4$Institute for Bioinformatics and Medical Informatics (IBMI), Eberhard-Karls University of Tübingen, Tübingen, Germany \\ 
	$^5$University of California, Berkeley, Berkeley, 94720, California, USA \\
	$^6$Institute of Computational Biology, Department of Computational Health, Helmholtz Munich, Germany \\ 
	$^7$Comprehensive Pneumology Center with the CPC-M bioArchive, Helmholtz Zentrum Munich, Member of the German Center for Lung Research (DZL), Munich, Germany \\ 
	$^8$TUM School of Life Sciences Weihenstephan, Technical University of Munich, Freising, Germany \\
	$^{9}$Department of Genetics, Genomics and Informatics, University of Tennessee Health Science Center, 71 S Manassas St, Memphis, 38163, Tennessee, USA \\ 
	$^{10}$Human Technopole, Viale Rita Levi-Montalcini 1, 20157, Milan, Italy \\
    }

@@

	Current pangenome graph construction methods exclude complex sequences or are reference-biased \citep{Chin2023, Minkin2016}.
	%, are tree-guided, or reference-based \citep{Li:2020, Hickey2023} leading to reference, order, or orientation bias. 
	%Although whole genome scaling approaches for unbiased pangenome graph construction \citep{Chin2023, Minkin2016} exist, their reliance on k-mer-based data structures often leads to unwanted complexity for downstream analysis. 
	One recent approach that overcomes such limitations is the PanGenome Graph Builder (PGGB) pipeline \citep{Garrison2023}. 
	PGGB iteratively refines an all-to-all whole-genome alignment graph that lets us explore sequence conservation and variation, infer phylogeny, and identify recombination events. 
	PGGB was already extensively evaluated \citep{Garrison2023, Andreace2023} and applied to build the first draft human pangenome reference \citep{Liao2023}.
	However, PGGB is implemented in bash: This (a) makes it difficult to deploy on HPC systems, (b) does not allow for a fine granular tuning of computing resources for different steps of the pipeline \citep{Sztuka2024}, and (c) limits its cluster scalability because PGGB can only use the resources of one node. These limitations greatly hinder the broad application of large-scale pangenomes.
	%SN: Given the high resource requirements of large-scale pangenomes, %these limitations are significantly limiting broad applications.

@@

	\vspace{-0.5cm}
